### 2016-07-12 conclusions.tex
More changes

\label{sec:conclusions}
OBsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this are, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed and how to guarantee replicability and scaling managing of the system. The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges. The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and administrable way. Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

>>>>

\label{sec:conclusions}
Obsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of the system. The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges. The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way. Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

### 2016-07-12 tools.tex
More changes

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary task execution in the \ac{Raspi}s. Given the
computing capabilities of the \ac{Raspi}, common computational demanding tasks
such as source code compilation and library linking can be challenginng if
not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
a description of how to cross-compile source code from the testbed
administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally in a \ac{Raspi} making
necessary to use a keyboard and a monitor.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to a lot of repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to and / or from multiple devices. To
simplify the management and avoid execution issues, \texttt{Fabric} provides
a Python library to ease the task of working with multiple devices from
a single \ac{PC}. The code below provides a small script that can perform
the few items above:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
<@\textcolor{gray}{
-------------------------------------}
\Reactivatenumber @>
from fabric.api import run, env, task
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

@@

The script functions can be called from the shell in the following
manner. \texttt{Fabric} will automatically search the directory for a file
named 'fabfile.py'. The syntax for executing a task and arguments is in the form
'fab <FUNCTION>:arg1,arg2,...:
\begin{lstlisting}[]
$ fab reboot
$ fab install:tmux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
The first function above reboots the \ac{Raspi}s in the lists of hosts, and
the install function installs a program given by an argument. In the above case,
\texttt{tmux} is installed.
%The two functions above reboots the \ac{Raspi}s specified in the list of hosts

Below is a slightly more complicated function call that copies a file from
a \ac{PC} to all \ac{Raspi}s in the list of hosts:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/hello_world",'~/'
[192.168.87.107] Executing task 'push'
[192.168.87.107] put: /home/<USER>/code/hello_world -> /home/pi/hello_world

@@

\texttt{Fabric} has many other useful functionalities that are useful in
controlling a bunch of \ac{Raspi}s.

%execute a script function by calling "fab install:feh"

\subsection{Long running jobs using SSH}

There may be times when a simulation has to run for days in the \ac{Raspi}s.
For this purpose, it may be inconvenient to keep a \ac{SSH} connection
open to the \ac{Raspi}s risking that the connection will be interrupted and
thus leading to the \ac{Raspi} automatically terminating the process.

%There may be times when it is desired to start a simulation that will
%run for days. For that purpose, it may not be desired to keep
%a \ac{SSH} connection open to the \ac{Raspi} risking that the connection
%will be interrupted and thus result in the \ac{Raspi} terminating
%the running simulation.

There exists a few methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly.
%There exists a few methods to avoid the risk of simulations terminating
%unexpectedly.
One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a \texttt{screen}
session which never terminates. Users can attach and detach from such a
session as desired.
%a session as it is
%desired. Programs within the screen session just keeps running.

The following procedure presents how to use \texttt{screen} with \ac{SSH}:
\begin{enumerate}
    \item login to a \ac{Raspi}
    \item open a screen session
    \item execute a command
    \item detach from the screen session
    \item terminate SSH connection
    \item login to the \ac{Raspi} again
    \item attach to screen session to see the program still running
\end{enumerate}

Start by establishing SSH connection to a \ac{Raspi} and open a screen session:
\begin{lstlisting}[]
$ SSH pi@<RASP_IP>
$ screen
Screen version 4.02.01 (GNU) 28-Apr-14
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Press space or return to clear the shell.

@@

%Problem when you logout, then applications will terminate. SCREEN is the answer.

\subsection{Cross-compilation: From your PC to the Raspberry Pi}
Some processing expensive tasks such as compiling code libraries can with
advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
This section will present how to setup a toolchain in a \ac{PC} that may
be used to cross compile C++ code to run in \ac{Raspi}s.
There will be given an example of compiling a simple C++ application to run
on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.

@@

Lets create a directory to hold our code:
\begin{lstlisting}[]
$ CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

>>>>

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary task execution in the \ac{Raspi}s. Given the
computing capabilities of the \ac{Raspi}, common computational demanding tasks
such as source code compilation and library linking can be challenginng if
not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
a description of how to cross-compile source code from the testbed
administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally in a \ac{Raspi} making
necessary to use a keyboard and a monitor. To do so, the testbed
administrator needs to put the memory cards in the \ac{Raspi}s and turn them
on, for them to be able to boot. The devices should boot properly and have
connectivity if the image and \ac{DHCP} server were properly set previously.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to many repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to and / or from multiple devices, etc. To
simplify the management and avoid execution issues, Fabric provides
a Python library to ease the major task of working with multiple devices from
a single \ac{PC}. First, the testbed administrator creates a directory to
hold the \texttt{Fabric} source code:

\begin{lstlisting}[]
$ export CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Then, the \texttt{\$CODEDIR/fabfile.py} file below provides a script with
some basic functionalities that can perform the few items above (i-iii).
In general, other administrators may require different functionalities
but that goes out of the scope of this work. Nevertheless, the following
file serves as a starting boilerplate:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
<@\textcolor{gray}{
-------------------------------------}
\Reactivatenumber @>
from fabric.api import env, task, sudo
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

@@

The previous \texttt{fabfile} shows three function that perform our
example tasks. These functions utilize variables and subsequent functions
from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
among others. Each of these \ac{API} functions permit to define environment
variables, create the administrator tasks through decorators or run the
mentioned task in \texttt{sudo} mode, respectively.
When a task is called from the terminal, Fabric searches the directory for
the \texttt{fabfile.py} file and executes the desired task. The syntax for
executing a task with arguments is in the form
\texttt{fab <FUNCTION>:arg1,arg2,...}. Some executions from the terminal
are shown as follows \todot{Test the following commands later and print the
outputs}:

\begin{lstlisting}[]
$ fab reboot
$ fab install:tmux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The first function above reboots the \ac{Raspi}s in the lists of hosts and
the second function installs a program given by an argument. In both cases,
after a successful \ac{SSH} connection is made, Fabric employs the
\ac{Raspi}'s \texttt{reboot} and \texttt{apt-get} commands in \texttt{sudo}
mode to do the required tasks. Below it is shown an example for
the \texttt{push} task which uses two arguments. Here we copy
\texttt{my\_file} from the testbed administrator \ac{PC} to all the host
\ac{Raspi}s:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/my_file",'~/'
[192.168.87.107] Executing task 'push'
[192.168.87.107] put: /home/<USER>/code/my_file -> /home/pi/my_file

@@

Fabric has many other functionalities that are useful in
controlling a large set of \ac{Raspi}s. For example, we may extract files
or run automated experiments. The included functionalities in
the \texttt{\$CODEDIR/fabfile.py} file will depend on the requirements
of the testbed administrator.

%execute a script function by calling "fab install:feh"
\subsection{Long running jobs using SSH}
There are times when a task may need to run for several hours or even days
in the \ac{Raspi}s, particularly when related to simulations or measurement
campaigns. For this purpose, it might be necessary to keep open a \ac{SSH}
connection in the \ac{Raspi}s without risking that the connection will
be interrupted and a given \ac{Raspi} terminating the task.

There are methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly. One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a shell
window, a \texttt{screen} session, which does not terminate even with
connectivity interruptions. Users can attach and detach from such a session
as desired. The following procedure presents how to use \texttt{screen} with
\ac{SSH} to: (i) Login to a \ac{Raspi}, (ii) open a screen session, (iii)
execute an example command, (iv) detach from the screen session, (v)
terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again
and (vii) attach to screen session to see the program still running.
From the testbed administrator, we start by establishing \ac{SSH} connection
to a \ac{Raspi} and open a screen session:
\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
$ screen
Screen version 4.02.01 (GNU) 28-Apr-14
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Press space or return to clear the shell.

@@

%Problem when you logout, then applications will terminate. SCREEN is the answer.
\subsection{Cross-compilation: From your PC to the Raspberry Pi}
Some processing expensive tasks such as compiling code libraries can with
advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
This section will present how to setup a toolchain in a \ac{PC} that may
be used to cross compile C++ code to run in \ac{Raspi}s.
There will be given an example of compiling a simple C++ application to run
on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.

@@
